---
title: "R Notebook"
output: html_notebook
---

Introduction to the Tydeverse

```{r}

library(gapminder)
library(dplyr)


gapminder

```

Filtering for one year

```{r}
gapminder %>%
  filter(year == 1957, country == "China")
```

Arranging observations by life expectancy
```{r}
gapminder %>%
  arrange(lifeExp)

#sort in descending order o lifeExp

gapminder %>%
  arrange(desc(lifeExp))
```

Filtering and arranging
```{r}
gapminder %>%
  filter(year == 1957) %>%
arrange(desc(pop))

```

Using mutate to change or create a colun

```{r}
#Use mutate to change lifeExp to be in months
gapminder %>%
  mutate(lifeExp=12*lifeExp)

#Use mutate to create a new column called lifeExpMonths
gapminder %>%
  mutate(lifeExpMonths = 12*lifeExp)

```

Combinging filter, mutate and arrange
```{r}
gapminder %>%
  filter(year == 2007) %>%
  dplyr::mutate(lifeExpMonths = 12* lifeExp) %>%
  arrange(desc(lifeExpMonths))  
```


Visualizing with ggplot2

```{r}

library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x=pop, y=gdpPercap)) + geom_point()
```


```{r}

ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) + geom_point()
```


```{r}

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) + 
  geom_point()+
  scale_x_log10()

```
Putting the x-and and y-axes on a log scale
```{r}

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x=pop, y=gdpPercap)) + 
  geom_point()+
  scale_x_log10()+
  scale_y_log10()
```

Adding color to scatter plot

```{r}
gapminder

ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color=continent)) + 
  geom_point()+
  scale_x_log10()

```

Adding size and color to a plot
```{r}

  ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color=continent, size=gdpPercap)) + 
    geom_point()+
    scale_x_log10()

```


Creating a subgraph for each continent
```{r}

  ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color=continent, size=gdpPercap)) + 
    geom_point()+
    scale_x_log10()+
  facet_wrap(~ continent)

```


Fetching by year
```{r}

  ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, color=continent, size=pop)) + 
    geom_point()+
    scale_x_log10()+
  facet_wrap(~ year)

```